---
title: "Activity Monitoring"
author: "Anirudha Mohanty"
date: "January 19, 2018"
output:
  html_document: default 
---

 
    Q1) Code for reading in the dataset and/or processing the data
    #Solution:
        a. Set working directory
        b. Download the file
        c. Unzip the file
        d. Analyse the file by using Summary,Str,head commands

```{r }

setwd("C:\\Coursera_Data\\Course5\\Course5_Week1\\Course5_Week1_Assignment")

download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", destfile="data.zip")
unzip (zipfile = "data.zip")


activityDs <- read.csv("activity.csv",header = TRUE)
activityDs$date <- as.Date(activityDs$date)
summary(activityDs)
str(activityDs)
head(activityDs)
```

    Q2.Histogram of the total number of steps taken each day
        #Solution:
            a) Find total number of steps taken on each day by using aggregate function.
            b)Remove the missing value from the above aggregate
            c)Use ggplot to plot the steps taken each day
        

```{r echo=TRUE}
library(ggplot2)

aggrStepsByDate <-aggregate(steps ~ date, activityDs, sum, na.rm = TRUE , na.action = na.omit)


ggplot(aggrStepsByDate, aes(x = steps))  + geom_histogram(fill = "blue", binwidth = 1000) +xlab("Number of Steps per Day")  +ylab("Number of times in a day(Count)")  + ggtitle("Total number of steps taken per day")  +
  theme(plot.title = element_text(hjust = 0.5))


``` 

    Q3.Mean and median number of steps taken each day 
        #Solution:
            *Use Mean & Median function
```{r}
mean(aggrStepsByDate$steps, na.rm=TRUE)

median(aggrStepsByDate$steps, na.rm=TRUE)

```


    Q4.Time series plot of the average number of steps taken
        #Solution:
            a) Find average number of steps taken on each inetrval (across all days) by using aggregate function.
            b)Remove the missing value from the above aggregate
            c)Use ggplot to plot the steps taken each day  
```{r echo=TRUE}
aggrStepsByInterval <-aggregate(steps ~ interval, activityDs, mean,na.rm = TRUE , na.action = na.omit)

ggplot(aggrStepsByInterval, aes(interval, steps)) + geom_line(color="blue") +labs(title="Average Daily Activity Pattern", x="Interval", y="Average steps")  +
  theme(plot.title = element_text(hjust = 0.5))

```

    Q5.The 5-minute interval that, on average, contains the maximum number of steps.
        #Solution:
            a) Find average number of steps taken on each inetrval (across all days) by using aggregate function and assign it to a variable.
            b)Remove the missing value from the above aggregate
            c)Find the maximum value by using which.ax function
        
```{r echo=TRUE}

mostSteps <- aggrStepsByInterval[which.max(aggrStepsByInterval$steps), ]

mostSteps
```

    Q6.Code to describe and show a strategy for imputing missing data
        #Solution:
            a)Calculate and report the total number of missing values in the dataset     (i.e. the total number of rows with NAs)
```{r}
missingNAN<- is.na(activityDs$steps)
totalMissing <- sum(as.numeric(missingNAN))
totalMissing
```
    
            b) Create new variable "steps_impute" .Fill missing values in  steps_impute with average steps taken during the same interval (across days). 
            c) use str function to see the structure of dataset.
```{r}
    for (i in 1:length(activityDs$steps)){
        if (is.na(activityDs$steps[i] == TRUE)){
           activityDs$steps_impute[i] <- aggrStepsByInterval$steps[match(activityDs$interval[i], aggrStepsByInterval$interval)]
            
        } else {
            activityDs$steps_impute[i] <- activityDs$steps[i]
        }
    }        
str(activityDs)

``` 

    Q7.Histogram of the total number of steps taken each day after missing values are imputed
        #Solution:
            a)Make a histogram of the total number of steps taken each day by using ggplot function. 
            
```{r echo=TRUE}

aggrImpStepsByDate <-aggregate(steps_impute ~ date, activityDs, sum)

str(aggrImpStepsByDate)

ggplot(aggrImpStepsByDate, aes(x=steps_impute))  + geom_histogram(fill = "green", binwidth =1000 ) +xlab("Number of (Imputed)Steps")  +ylab("Number of times in a day")  + ggtitle("Steps taken per day")  +
  theme(plot.title = element_text(hjust = 0.5))
```
            
            b)Use mean/median function to calculate mean and median of total number of steps taken per day
   

```{r}
mean(aggrStepsByDate$steps)

median(aggrStepsByDate$steps)

```

    Q8)Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
    #Solution:
        a) Create "day" variable in the dataset using weekday function
        b) Use conditional operator to get Weekday/Weekend flag
        c) Aggregate total steps taken for all the interval during weekday and weekend
        d) Use ggplot function to show steps taken during each interval
        facet_wrap annotation can be use to split the graph based on weekday and weekend
        
```{r}
 activityDs$day <- weekdays(activityDs$date,abbreviate = TRUE)    
    activityDs$weekdayFlag <- ifelse(activityDs$day == "Sun","Weekend",ifelse(activityDs$day == "Sat","Weekend","Weekday"))
    aggrStepWeekdayInt <- aggregate(steps_impute ~ interval +weekdayFlag , activityDs, sum)
    
    ggplot(aggrStepWeekdayInt, aes(interval,steps_impute)) + geom_line(color="blue")+ facet_wrap(~weekdayFlag, nrow = 2,ncol = 1) +labs(x="Interval", y="Number of steps") 
    
```
